

```{r}
library(tidyverse)
library(dplyr)
```




```{r}
gdp_df <- read.csv("../data/gdp_per_capita.csv")
```
```{r}
  tail(gdp_df,n=10)
```

```{r}
head(gdp_df, n=10)
```

```{r}
gdp_df
```

```{r}
# hello testing
gdp_df <- gdp_df |> 
  select(-'Value.Footnotes')
```

```{r}
#  #3
gdp_df |> 
  count(Year) |> 
  arrange(desc(Year))
```

```{r}
year_counts <- gdp_df |> 
  count(Year) |> 
  arrange(Year)
```
The years go from 1990 -> 2022, and the graph will show how many observations there are for each year. 

```{r}
ggplot(year_counts, aes(x = Year, y = n)) + geom_point()
```
#4  How many countries are represented in this dataset? Which countries have the fewest observations?

```{r}
country_counts <- gdp_df |> 
  count(Country.or.Area) |> 
  arrange(n)
```

```{r}
country_counts
```
There are 242 countries in the dataframe, and Djibouti has the fewest observations. (We know this because we did arrange and didn't say desc, so it is ascending.)


#5 Create a new tibble by subsetting `gdp_df` to just the year 2021. Call this new tibble `gdp_2021`.

```{r}
gdp_2021 <- gdp_df |> 
  filter(Year == 2021)
```


```{r}
gdp_2021
```

#6 Use `summary()` to find the summary statistics for GDP per capita in 2021.


```{r}
summary(gdp_2021)
```
#7  Create a histogram of GDP Per Capita numbers for 2021 (you may wish to adjust the number of bins for your histogram).


```{r}
ggplot(gdp_2021, aes(x = Value)) + geom_histogram(bins = 25)
```

#8 Find the top 5 counties and bottom 5 countries by GDP per capita in 2021.

  
```{r}

gdp_2021 <- gdp_2021  |> 
  arrange(desc(Value))
```
  
```{r}
head(gdp_2021, n=5)
```
```{r}
tail(gdp_2021, n=5)
```
  The top 5 were Luxembourg, Singapore, Ireland, Qatar, Bermuda
  The bottom 5 were Niger, Somalia, Dem. Rep. Congo, Central African Republic, and Burundi.
  
  #9 Now, return to the full dataset, `gdp_df`. Pivot the data for 1990 and 2021 (using `pivot_wider()` function) so that each row corresponds to a country, each column corresponds to a year, and the values in the table give the GDP_Per_Capita amount. Drop any rows that are missing values for either 1990 or 2021. Save the result to a tibble named `gdp_pivoted`.

  
```{r}
```


```{r}
gdp_pivoted <- gdp_df |> 
  filter(Year == 2021 | Year == 1990) |> 
  pivot_wider(names_from = Year, values_from = Value) |> 
  drop_na()
```
  
  
```{r}
gdp_pivoted
```

#10  Create a new column in `gdp_pivoted` named `Percent_Change`. This column should contain the percent change in GDP_Per_Capita from 1990 to 2021. Hint: Percent change is calculated as 100*(New Value - Old Value) / Old Value.


```{r}

```



```{r}
gdp_pivoted
```


```{r}
gdp_pivoted <- gdp_pivoted |> 
  mutate(Percent_Change = (100 * (`2021` - `1990`)) / `1990`)
```

```{r}
gdp_pivoted
```

#11 How many countries experienced a negative percent change in GDP per capita from 1990 to 2021?


  
```{r}
gdp_pivoted |> 
  filter(Percent_Change < 0)
```
  19 countries experiences a negative percent change from 1990 to 2021
  
  
  #12 Which country had the highest % change in GDP per capita? Create a line plot showing these country's GDP per capita for all years for which you have data. Put both line charts on the same plot.
  
```{r}
gdp_pivoted |> 
  arrange(desc(Percent_Change))
          
```
  Equatorial Guinea has the largest percent change
  
```{r}
gdp_df_top <- gdp_df |> 
  filter(Country.or.Area == 'Equatorial Guinea' | Country.or.Area == 'China')
```
  
```{r}
ggplot(gdp_df_top, aes(x = Year, y = Value, color = Country.or.Area)) + geom_line()
```

#13  Read in continents.csv contained in the `data` folder into a new tibble called `continents`. We will be using this tibble to add a new column to our dataset.

```{r}
continents <- read.csv("../data/continents.csv")
```

#14 Merge gdp_df and continents. Keep only the countries that appear in both data frames. Save the result back to gdp_df.


```{r}
gdp_df <- gdp_df |> 
  rename(Country = Country.or.Area)
```





```{r}
gdp_df <- inner_join(continents, gdp_df, by = 'Country')
```

#15 Determine the number of countries per continent. Create a bar chart showing this.


```{r}
gdp_df_countries <- gdp_df |> 
  group_by(Continent) |> 
  summarize(CountryCount = n_distinct(Country))
```

```{r}
ggplot(gdp_df_countries, aes(x = Continent, y = CountryCount)) + geom_col()
```

#16 Create a boxplot showing GDP per capita in 2021 split out by continent. What do you notice?
```{r}
gdp_2021 <- gdp_df |> 
  filter(Year == 2021)

gdp_2021
```

```{r}
ggplot(gdp_2021, aes(x = Value, y = Continent)) + geom_boxplot()
```
I notice that Europe has the highest median gdp_per_capita value and the highest outliers, while Africa has the lowest median.

#17 Read life_expectancy.csv into a tibble named life_expectancy. Do not modify the csv file in order to read this data in. 

```{r}
life_expectancy <- read.csv("../data/life_expectancy.csv", skip = 3)
life_expectancy
```
#18 Drop the Country Code, Indicator Name, and Indicator Code columns. Then use `pivot_longer` to convert your data from wide to long. That is, instead of having one row per country and multiple colums per year, we want to have multiple rows per country and a single column for year. After melting, rename the columns to `Country`, `Year`, and `Life_Expectancy`.
```{r}
life_expectancy <- life_expectancy |> 
  select(-Country.Code, -Indicator.Name, -Indicator.Code) |> 
  pivot_longer(cols = -Country.Name, names_to = "Year", values_to = "Life_Expectancy") |> 
  rename(Country = Country.Name) |> 
  mutate(Year = as.integer(str_remove(Year, "^X")))
  

life_expectancy
```
#19 What was the first country with a life expectancy to exceed 80?


```{r}
life_expectancy |> 
  filter(Life_Expectancy >= 80) |> 
  arrange(Year)
```
It was Japan in 1996 that was the first country to exceed a life expectancy of 80 years.


#20 Merge `gdp_df` and `life_expectancy`, keeping all countries and years that appear in both tibbles. Save the result to a new tibble named `gdp_le`. If you get any errors in doing this, read them carefully and correct them. Look at the first five rows of your new data frame to confirm it merged correctly. Also, check the last five rows to make sure the data is clean and as expected.

```{r}
gdp_le <- inner_join(gdp_df, life_expectancy, by = c("Country", "Year"))
gdp_le
```

#21 Create a new tibble, named `gdp_le_2021` by extracting data for the year 2021 from `gdp_le`. How many countries have a life expectancy of at least 80 in 2021?

```{r}
gdp_le_2021 <- gdp_le |> 
  filter(Year == 2021)

gdp_le_2021 |> 
  filter(Life_Expectancy >= 80)
```

31 countries had atleast a life expectancy of 80 in the year of 2021

#22 Find the countries that had the top 3 largest GDP per capita figures for 2021. Create a plot showing the change in life expectancy over time for these three countries. This plot should be faceted so that each country is contained in its own figure.

```{r}
gdp_le_2021 |> 
  arrange(desc(Value))

gdp_life_exp_top3 <- gdp_le |> 
  filter(Country == "Luxembourg" | Country == "Singapore" | Country == "Ireland")

gdp_life_exp_top3

ggplot(gdp_life_exp_top3, aes(x = Year, y = Life_Expectancy)) + geom_line() + facet_wrap(~Country)
```

#23 Create a scatter plot of Life Expectancy vs GDP per Capita for the year 2021. What do you notice?

```{r}
ggplot(gdp_le_2021, aes(x = Life_Expectancy, y = Value)) + geom_point()
```
The GDP_Per_Capita seems to go up exponentially with the life expectancy going up. It could also be said the same way vice versa. 

#24 Find the correlation between Life Expectancy and GDP per Capita for the year 2021. What is the meaning of this number?

```{r}
install.packages("corrr")
library(corrr)

correlate(gdp_le_2021 |> select(Life_Expectancy, Value))
```
The correlation value we got was .745, which means it correlates decently well with 1 being perfectly.

#25  Add a column to `gdp_le_2021` and calculate the logarithm of GDP per capita. Find the correlation between the log of GDP per capita and life expectancy. How does this compare to the calculation in the previous part? Look at a scatter plot to see if the result of this calculation makes sense.

```{r}
gdp_le_2021 <- gdp_le_2021 |> 
  mutate(gdp_log = log(Value))

gdp_le_2021
```

```{r}
correlate(gdp_le_2021 |> select(Life_Expectancy, gdp_log))
```
The correlation factor is better in this one than the previous, with it being closer to 1.

```{r}
ggplot(gdp_le_2021, aes(x = Life_Expectancy, y = gdp_log)) + geom_point()
```
this looks like a more linear line now than before.



