

```{r}
library(tidyverse)
library(dplyr)
```




```{r}
gdp_df <- read.csv("../data/gdp_per_capita.csv")
```
```{r}
  tail(gdp_df,n=10)
```

```{r}
head(gdp_df, n=10)
```

```{r}
gdp_df
```

```{r}
# hello testing
gdp_df <- gdp_df |> 
  select(-'Value.Footnotes')
```

```{r}
#  #3
gdp_df |> 
  count(Year) |> 
  arrange(desc(Year))
```

```{r}
year_counts <- gdp_df |> 
  count(Year) |> 
  arrange(Year)
```
The years go from 1990 -> 2022, and the graph will show how many observations there are for each year. 

```{r}
ggplot(year_counts, aes(x = Year, y = n)) + geom_point()
```
#4 

```{r}
country_counts <- gdp_df |> 
  count(Country.or.Area) |> 
  arrange(n)
```

```{r}
country_counts
```
There are 242 countries in the dataframe, and Djibouti has the fewest observations. (We know this because we did arrange and didn't say desc, so it is ascending.)


#5

```{r}
gdp_2021 <- gdp_df |> 
  filter(Year == 2021)
```


```{r}
gdp_2021
```

#6 


```{r}
summary(gdp_2021)
```
#7


```{r}
ggplot(gdp_2021, aes(x = Value)) + geom_histogram(bins = 25)
```

#8

  
```{r}

gdp_2021 <- gdp_2021  |> 
  arrange(desc(Value))
```
  
```{r}
head(gdp_2021, n=5)
```
```{r}
tail(gdp_2021, n=5)
```
  The top 5 were Luxembourg, Singapore, Ireland, Qatar, Bermuda
  The bottom 5 were Niger, Somalia, Dem. Rep. Congo, Central African Republic, and Burundi.
  
  #9
  
```{r}
```


```{r}
gdp_pivoted <- gdp_df |> 
  filter(Year == 2021 | Year == 1990) |> 
  pivot_wider(names_from = Year, values_from = Value) |> 
  drop_na()
```
  
  
```{r}
gdp_pivoted
```

#10


```{r}

```



```{r}
gdp_pivoted
```


```{r}
gdp_pivoted <- gdp_pivoted |> 
  mutate(Percent_Change = (100 * (`2021` - `1990`)) / `1990`)
```

```{r}
gdp_pivoted
```

#11


  
```{r}
gdp_pivoted |> 
  filter(Percent_Change < 0)
```
  19 countries experiences a negative percent change from 1990 to 2021
  
  
  #12
  
```{r}
gdp_pivoted |> 
  arrange(desc(Percent_Change))
          
```
  Equatorial Guinea has the largest percent change
  
```{r}
gdp_df_top <- gdp_df |> 
  filter(Country.or.Area == 'Equatorial Guinea' | Country.or.Area == 'China')
```
  
```{r}
ggplot(gdp_df_top, aes(x = Year, y = Value, color = Country.or.Area)) + geom_line()
```

#13 

```{r}
continents <- read.csv("../data/continents.csv")
```

#14


```{r}
gdp_df <- gdp_df |> 
  rename(Country = Country.or.Area)
```





```{r}
gdp_df <- inner_join(continents, gdp_df, by = 'Country')
```

#15 Determine number of countries per continent, do a bar graph of it.


```{r}
gdp_df_countries <- gdp_df |> 
  group_by(Continent) |> 
  summarize(CountryCount = n_distinct(Country))
```

```{r}
ggplot(gdp_df_countries, aes(x = Continent, y = CountryCount)) + geom_col()
```

#16
```{r}
gdp_2021 <- gdp_df |> 
  filter(Year == 2021)

gdp_2021
```

```{r}
ggplot(gdp_2021, aes(x = Value, y = Continent)) + geom_boxplot()
```
I notice that Europe has the highest median gdp_per_capita value and the highest outliers, while Africa has the lowest median.

#17

```{r}
life_expectancy <- read.csv("../data/life_expectancy.csv", skip = 3)
life_expectancy
```
#18
```{r}
life_expectancy <- life_expectancy |> 
  select(-Country.Code, -Indicator.Name, -Indicator.Code) |> 
  pivot_longer(cols = -Country.Name, names_to = "Year", values_to = "Life_Expectancy") |> 
  rename(Country = Country.Name) |> 
  mutate(Year = as.integer(str_remove(Year, "^X")))
  

life_expectancy
```
#19


```{r}
life_expectancy |> 
  filter(Life_Expectancy >= 80) |> 
  arrange(Year)
```
It was Japan in 1996 that was the first country to exceed a life expectancy of 80 years.


#20

```{r}
gdp_le <- inner_join(gdp_df, life_expectancy, by = c("Country", "Year"))
gdp_le
```

#21

```{r}
gdp_le_2021 <- gdp_le |> 
  filter(Year == 2021)

gdp_le_2021 |> 
  filter(Life_Expectancy >= 80)
```

31 countries had atleast a life expectancy of 80 in the year of 2021

#22

```{r}
gdp_le_2021 |> 
  arrange(desc(Value))

gdp_life_exp_top3 <- gdp_le |> 
  filter(Country == "Luxembourg" | Country == "Singapore" | Country == "Ireland")

gdp_life_exp_top3

ggplot(gdp_life_exp_top3, aes(x = Year, y = Life_Expectancy)) + geom_line() + facet_wrap(~Country)
```

#23

```{r}
ggplot(gdp_le_2021, aes(x = Life_Expectancy, y = Value)) + geom_point()
```
The GDP_Per_Capita seems to go up exponentially with the life expectancy going up. It could also be said the same way vice versa. 

#24

```{r}
install.packages("corrr")
library(corrr)

correlate(gdp_le_2021 |> select(Life_Expectancy, Value))
```
The correlation value we got was .745, which means it correlates decently well with 1 being perfectly.

#25

```{r}
gdp_le_2021 <- gdp_le_2021 |> 
  mutate(gdp_log = log(Value))

gdp_le_2021
```

```{r}
correlate(gdp_le_2021 |> select(Life_Expectancy, gdp_log))
```
The correlation factor is better in this one than the previous, with it being closer to 1.

```{r}
ggplot(gdp_le_2021, aes(x = Life_Expectancy, y = gdp_log)) + geom_point()
```
this looks like a more linear line now than before.



