

```{r}
library(tidyverse)
```




```{r}
gdp_df <- read.csv("../data/gdp_per_capita.csv")
```
```{r}
  tail(gdp_df,n=10)
```

```{r}
head(gdp_df, n=10)
```

```{r}
gdp_df
```

```{r}
# hello testing
gdp_df <- gdp_df |> 
  select(-'Value.Footnotes')
```

```{r}
#  #3
gdp_df |> 
  count(Year) |> 
  arrange(desc(Year))
```

```{r}
year_counts <- gdp_df |> 
  count(Year) |> 
  arrange(Year)
```
The years go from 1990 -> 2022, and the graph will show how many observations there are for each year. 

```{r}
ggplot(year_counts, aes(x = Year, y = n)) + geom_point()
```
#4 

```{r}
country_counts <- gdp_df |> 
  count(Country.or.Area) |> 
  arrange(n)
```

```{r}
country_counts
```
There are 242 countries in the dataframe, and Djibouti has the fewest observations. (We know this because we did arrange and didn't say desc, so it is ascending.)


#5

```{r}
gdp_2021 <- gdp_df |> 
  filter(Year == 2021)
```


```{r}
gdp_2021
```

#6 


```{r}
summary(gdp_2021)
```
#7


```{r}
ggplot(gdp_2021, aes(x = Value)) + geom_histogram(bins = 25)
```

#8

  
```{r}

gdp_2021 <- gdp_2021  |> 
  arrange(desc(Value))
```
  
```{r}
head(gdp_2021, n=5)
```
```{r}
tail(gdp_2021, n=5)
```
  The top 5 were Luxembourg, Singapore, Ireland, Qatar, Bermuda
  The bottom 5 were Niger, Somalia, Dem. Rep. Congo, Central African Republic, and Burundi.
  
  #9
  
```{r}
```


```{r}
gdp_pivoted <- gdp_df |> 
  filter(Year == 2021 | Year == 1990) |> 
  pivot_wider(names_from = Year, values_from = Value) |> 
  drop_na()
```
  
  
```{r}
gdp_pivoted
```

#10


```{r}

```



```{r}
gdp_pivoted
```


```{r}
gdp_pivoted <- gdp_pivoted |> 
  mutate(Percent_Change = (100 * (`2021` - `1990`)) / `1990`)
```

```{r}
gdp_pivoted
```

  